Testing and Isolation Efficacy: Insights from a Simple Epidemic Model

### 2021-05-04 SIR_manuscript.tex
way too much figure tweaking
adding proof of decreasing R0 wrt thetas'

\begin{figure}[h!]
\centering
\begin{subfigure}[t]{.45\textwidth}
\centering
\includegraphics[width=\linewidth]{codes/R0contour_random.pdf}
\caption{}\label{p.a}
\end{subfigure}
%
\begin{subfigure}[t]{.45\textwidth}
\centering
\includegraphics[width=\linewidth]{codes/R0contour_TTI.pdf}
\caption{}\label{p.b}
\end{subfigure}
\caption{
{\bf A comparison of the behaviour of the basic reproduction number, $\Rnum$, between random versus targeted testing strategies at different levels of testing and isolation.}
We numerically evaluate $\Delta$ \eqref{R0}, reflecting the reduction of $\Rnum$ with respect to testing and isolation. We use the following parameters (see Table \ref{tab:params}):
$N_0=1 \times 10^6$, $\omega \in [0.1,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.01]$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.339$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$. Contours of $\Delta$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$. The black solid line in each panel represents the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, i.e., the $\Delta$ corresponding to the threshold of $\Rnum=1$. \bmb{we still have some collisions in the y-axis tick labels}
}
\label{pan}
\end{figure}

@@

This is to provide a mathematical materials to prove that $\pro \Delta$ can be positive or negative, with $\Delta$ defined in Eq. \eqref{eq:abc}, and thus $\pro \Rnum < 0$, where $\Rnum$ is the basic reproduction number, given in Eqs.\,\eqref{R0}. We can rewrite matrix $G_{11}$ in \eqref{G11} in the following form to simplify the calculations.
\begin{equation}
\label{G112}
G_{11} = \frac{\beta}{\gamma \, C} 
\left[ \begin {array}{c}  (\omega-\rho)/\omega \\ (1-\theta\_w)\rho/\omega  \end {array} \right]
\left[ \begin {array}{cccc} 
C-C_1, C-C_1-C_2\end {array} \right],
\end{equation}
where $C$ is the same as the one in Eq. \eqref{eq:abc}, i.e., \ali{all equations after this point need to be updated.}
$$C=(\omega+\gamma)(\gamma(\omega+\gamma)+(\omega\,p_I+\gamma)\,\hat F_I),$$
and $C_1$ and $C_2$ are 
\begin{align}
\label{eq:C12}
C_1 =& (\omega+\gamma)(\theta\_w \, \gamma+\theta\_c\, \omega\, p_I) \hat F_I,\\
C_2 =& (\omega+\gamma)\gamma^2\,\theta\_w+(\theta\_w-\theta\_c)\gamma\,\omega\,p_I\,\hat F_I,
\end{align}
where $\hat F_I$ is given in Eq. \eqref{eq:fi}.
$\Rnum$ is in the same form as in Eq. \eqref{R0}  
$$\Rnum= \frac{\beta}{\gamma} (1-\Delta).$$

>>>>

\begin{figure}[h!]
\centering
\begin{subfigure}[t]{.45\textwidth}
\centering
\includegraphics[width=\linewidth]{codes/R0contour_random.pdf}
\caption{}\label{p.a}
\end{subfigure}
%
\begin{subfigure}[t]{.45\textwidth}
\centering
\includegraphics[width=\linewidth]{codes/R0contour_TTI.pdf}
\caption{}\label{p.b}
\end{subfigure}
\caption{
{\bf A comparison of the behaviour of the basic reproduction number, $\Rnum$, between random versus targeted testing strategies at different levels of testing and isolation.}
We numerically evaluate $\Delta$ \eqref{R0}, reflecting the reduction of $\Rnum$ with respect to testing and isolation. We use the following parameters (see Table \ref{tab:params}):
$N_0=1 \times 10^6$, $\omega \in [0.1,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.01]$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.339$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$. Contours of $\Delta$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$. The black dotted line in each panel represents the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, i.e., the $\Delta$ corresponding to the threshold of $\Rnum=1$.
}
\label{pan}
\end{figure}

@@

This is to provide a mathematical materials to prove that $\pro \Delta$ can be positive or negative, with $\Delta$ defined in Eq. \eqref{eq:abc}, and thus $\pro \Rnum < 0$, where $\Rnum$ is the basic reproduction number, given in Eqs.\,\eqref{R0}. We can rewrite matrix $G_{11}$ in \eqref{G11} in the following form to simplify the calculations.
\begin{equation}
\label{G112}
G_{11} = \frac{\beta}{\gamma \, C} 
\left[ \begin {array}{c}  (\omega-\rho)/\omega \\ (1-\theta\_w)\rho/\omega  \end {array} \right]
\left[ \begin {array}{cccc} 
C-C_1, C-C_1-C_2\end {array} \right],
\end{equation}
where $C$ is the same as the one in Eq. \eqref{eq:abc}, i.e., 
$$C=(\omega+\gamma)(\gamma(\omega+\gamma)+(\omega\,p_I+\gamma)\,\hat F_I),$$
and $C_1$ and $C_2$ are 
\begin{align}
\label{eq:C12}
C_1 =& (\omega+\gamma)(\theta\_w \, \gamma+\theta\_c\, \omega\, p_I) \hat F_I,\\
C_2 =& (\omega+\gamma)\gamma^2\,\theta\_w+(\theta\_w-\theta\_c)\gamma\,\omega\,p_I\,\hat F_I,
\end{align}
where $\hat F_I$ is given in Eq. \eqref{eq:fi}.
$\Rnum$ is in the same form as in Eq. \eqref{R0}  
$$\Rnum= \frac{\beta}{\gamma} (1-\Delta),$$
where 
\begin{equation}
\label{eq:del4}
\Delta= \frac{1}{N_0\,C}\big(C_1\,S\_u^*+((C_1+C_2)(1-\theta\_w)+C\,\theta\_w)\,S\_n^*\big).
\end{equation}
Note that for analysis convention, we let $N_0=1$, thus $S\_u^*$ and $S\_n^*$ are in the scale of 0 to 1. 
To derive the first result, i.e., how changes in $\theta\_w$ and $\theta\_c$ affects $\Rnum$, we factor these parameters in $\Delta$ \eqref{eq:del4}. Thus, $\Delta$ can be rewritten as
\begin{equation}
\label{eq:del4_theta}
\Delta= \frac{1}{C}\Big(
-D_1\,S\_n^*\,\boldsymbol {\theta\_w^2}+
\big(-\omega^2\,p_I\,\hat F_I\,S\_n^*\,\boldsymbol{\theta\_c}\,+\,D_2\,S\_n^*\,+\,\gamma\,\hat F_I(\omega+\gamma) \big)\boldsymbol{\theta\_w}+
(\omega+\gamma\,S\_u^*)\omega\,p_I\,\hat F_I\,\boldsymbol{\theta\_c}
\Big),
\end{equation}
where
\begin{align}
\label{eq:del4_d1d2}
D_1=& (\omega+\gamma)\gamma^2+(\omega+\gamma+\omega\,p_I)\gamma\,\hat F_I, \\
D_2=& (3\omega+2\gamma)\gamma^2+(\omega+\gamma+2\omega\,p_I)\gamma\,\hat F_I+(\gamma+p_I\,\hat F_I)\omega^2.
\end{align}
The scaled expression of $\Delta$, specifically $\Delta\,C$, in Eq.\,\eqref{eq:del4_theta} is linear in $\theta\_c$ with a positive coefficient given by 
$$(\gamma\,S\_u^*\,+\omega(1-\theta\_w\,S\_u^*))\omega\,p_I\,\hat F_I.$$
This results in increasing $\Delta$, thus decreasing $\Rnum$ with respect to $\theta\_c$, that is $\pder[\Rnum]{\theta\_c}\leq 0$. Note that $C$ is independant of $\theta\_c$ and $\theta\_w$.

With a similar logic, the expression of $\Delta\,C$ derived from Eq.\,\eqref{eq:del4_theta} is a concave-down quadratic equation in $\theta\_w$, given by
\begin{equation}
\label{eq:del4_thetaw}
-D_1\,S\_n^*\,\boldsymbol {\theta\_w^2}+
\big(-\omega^2\,p_I\,\hat F_I\,S\_n^*\,\boldsymbol{\theta\_c}\,+\,D_2\,S\_n^*\,+\,\gamma\,\hat F_I(\omega+\gamma) \big)\boldsymbol{\theta\_w}.
\end{equation}
We show that the feasible range of $\theta\_w$ lies btween 0 and the vertex of this parabola where the parabola is increasing in $\theta\_w$, and so does $\Delta$ which results in inferring $\pder[\Rnum]{\theta\_w}\leq 0$.
It is enough to show that partial derivative of the expression \eqref{eq:del4_thetaw} with respect to $\theta\_w$ at $\theta\_w=1$ is non-negative. It follows that
\begin{align}
\label{eq:d_del4_thetaw}
\pder[\Delta\,C]{\theta\_w}\bigg\rvert_{\theta\_w=1} =&
(D_2-2D_1-\omega^2\,p_I\,\hat F_I\,\theta\_c)\,S\_n^*+\gamma\,\hat F_I(\omega+\gamma) \\
=& (\gamma(\omega+\gamma)+\gamma\omega^2+(1-\theta\_c)\omega^2\,p_I\,\hat F_I )\,S\_n^*
+\gamma(\omega+\gamma)\,\hat F_I\,(1-S\_n^*),
\end{align}
which is a positive quantity, given that $\theta\_c$ and $S\_n^*$ vary between 0 and 1.

\ali{stoped here!}
